Molecular dynamics simulation of synchronization in driven particles

### 2021-01-14 manuscript/manuscript_ajp_2020.tex
moved multiparticle to conclusion

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion.
If $A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima
with no net velocity.  
%In Fig.~\ref{fig:0}

%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%The average velocity $\bar{v}_y$
%is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
%over the period of the driving force.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1

The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and length of the particle hops
forward and backward in the landscape.
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
We explore this in Ex.~\ref{ex:backwards}.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$,
as shown in Fig.~\ref{fig:1}(b).
We keep $F^{dc}$ constant
for $\Delta t = 10^4$ timesteps
and 
sample the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$.

Synchronization
of the driving force with the landscape,
despite the overall magnitude
of $F^{dc}$.

The particle 
achieves a variety of oscillation modes as a function of $F^{dc}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\bar{v}_{y}$
over a range of driving forces $F^{dc}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}(b).
Here $\langle v_{y} \rangle$ is increasing in non-uniform steps,
with a quantized height of
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.  
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{(Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$.  We fix the amplitude of $F^d(t)$ as $F^{ac}=0.05$ as in Fig.~\ref{fig:0}}.
\label{fig:1}
\end{figure}
\end{center}

>>>>

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion.
If $A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima
with no net velocity.  
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1

The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and length of the particle motions 
forward and backward in the landscape.
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
We explore this in Ex.~\ref{ex:backwards}.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$,
as shown in Fig.~\ref{fig:1}(b).
We keep $F^{dc}$ constant
for $\Delta t = 10^5$ simulation timesteps
and 
sample the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$.
We plot the results in Fig.~\ref{fig:1}
for $F^{ac} = 0.05$, $f=0.01$, and the landscape force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
The steps in $\langle v_y \rangle$ occur because 
the average velocity of the particle
is constant across a range of $F^{dc}$.
This synchronization
of the particle velocity can be measured
in terms of the landscape period.
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle.
The average velocity $\langle{v}_y\rangle$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
At low $F^{dc}$, the particle starts and ends in the same minima.

The particle 
achieves a variety of oscillation modes as a function of $F^{dc}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\bar{v}_{y}$
over a range of driving forces $F^{dc}$.
We illustrate mode-locking in 
the velocity-force plot in Fig.~\ref{fig:1}.
Here $\langle v_{y} \rangle$ is increasing in non-uniform steps,
with a quantized height of
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.  
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$.
  We fix the remaining parameters in $F^d(t)$ as $F^{ac}=0.05$ and $f = 0.01$ as in Fig.~\ref{fig:0}}.
\label{fig:1}
\end{figure}
\end{center}
